---
title: "DSAA811 - Final Presentation"
author: Sharon Van Den Berg
date: "`r Sys.Date()`"
output: 
  ioslides_presentation:
      css: style.css
      logo: ../images/UOW.png
      self_contained: false
      widescreen: true
      theme: yeti
header-includes:
  - \usepackage{titling}
  - \usepackage{setspace}
  - \onehalfspacing
runtime: shiny

bibliography: DSAA811.bib
link-citations: TRUE
citation_package: apa-annotated.csl
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=F,message=F,cache=F)
library(knitr)
library(tinytex)
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(forcats)
library(lme4)
library(lmerTest)
library(sf)
library(kableExtra)
library(sjmisc)
library(maps)
library(magrittr)
library(shiny)
library(gganimate)
library(DT)
library(gapminder)
```

```{r loadingData, include=FALSE}
# copy the R logo to the current directory, this gives the image a local path
file.copy(file.path(R.home("home"), ".", "Rings.png"), '.')
```

```{r child = '00_loader.Rmd' , eval = TRUE}
```

```{r setupSettings, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = F,
                      message = F,
                      fig.width = 6.5,
                      fig.height = 5.5,
                      cache = F) 
```

# Research Questions

## What are we trying to discover

*1) Have the events of the Olympics stayed the same? What way are they different today as apposed to the inception date?*

*2) Is the number of medals obtained proportionate to the number of athletes? Can we send one athlete and have them obtain twice as many medals in events as an individual athlete? Is the number of athletes on a team proportionate to the number of medals that the team can win?*

*3) What are the sports that have the most events and as such the highest potential to win a medal? Has this always been the case? What are the future predictions for medal obtainment moving into the next Olympic games in both Summer and Winter Olympics?*

# Data Description

## Data Source

- Olympic Image sourced from [@a2025_what,]

- [@bansal_2021_olympics_] data set called "Olympics\_" was compiled by "Harsh Bansal"

- [@a262588213843476_2021_countries] was sourced from Github containing the geoloications that match with the NOC codes

## Olympic Team Map
```{r}
  Regions_all <- factor(labels = c("Summer", "Winter","Everything"),
                        levels = c("Summer", "Winter","Everything"))
```


```{r warnings = FALSE, echo = FALSE}
shinyApp(
  ui = fluidPage(
  "Athlete locations from across the world.",
    selectInput("region", "Olympics:",
                choices = levels(Regions_all), selected = "Everything"),
    plotOutput("WorldPlot")
  ),
  server = function(input,output) {
    output$WorldPlot = renderPlot({ 
      drawMap(input$region, "All Countries that have competed at the selected olympic games")
    })
  } 
)
```

## Number of Medal Winners per Sport

```{r}
  SummerOrWinter <- factor(labels = c("Summer", "Winter"),
                        levels = c("Summer", "Winter"))
```

```{r warnings = FALSE, echo = FALSE}
shinyApp(
  ui = fluidPage(
  "Choose the Olympics and Year",
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Olympics:",
                    choices = levels(SummerOrWinter), selected = "Summer"),
      selectInput("nYear", "Olympic Year:",
                  choices = c(1896), selected  = 1986)),
      mainPanel(tableOutput("Winners")))
    ),
  server = function(input,output, session) {
    observeEvent(input$region, {
      if (input$region == "Summer") {
        choices <- events %>% filter(Season == "Summer") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      } else {
        choices <- events %>% filter(Season == "Winter") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      }
    updateSelectInput(session, "nYear", choices = choices)
    })
    output$Winners = renderTable(MedalWinners(input$region, input$nYear))
  } 
)
```

## Number of Medal Winners Per Team and Event Entries

```{r warnings = FALSE, echo = FALSE}
shinyApp(
  ui = fluidPage(
  "Choose the Olympics and Year",
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Olympics:",
                    choices = levels(SummerOrWinter), selected = "Summer"),
      selectInput("nYear", "Olympic Year:",
                  choices = c(1896), selected  = 1896)),
      mainPanel(tableOutput("Winners")))
    ),
  server = function(input,output, session) {
    observeEvent(input$region, {
      if (input$region == "Summer") {
        choices <- events %>% filter(Season == "Summer") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      } else {
        choices <- events %>% filter(Season == "Winter") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      }
    updateSelectInput(session, "nYear", choices = choices)
    })
    output$Winners = renderTable(MedalBreakDownByTeam(input$region, input$nYear) %>% select(Team, Gold, Silver, Bronze, Events = TotalEvents, Medals = TotalMedals))
  } 
)
```

## The proportion of Medals Per Team Entries

```{r warnings = FALSE, echo = FALSE}
shinyApp(
  ui = fluidPage(
  "Choose the Olympics and Year",
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Olympics:",
                    choices = levels(SummerOrWinter), selected = "Summer"),
      selectInput("nYear", "Olympic Year:",
                  choices = c(1896), selected  = 1896),
      radioButtons("sType", "Bar Type:",
                   choices = c("Fill", "Dodge"), selected = "Fill"),
      sliderInput("nCountry", "Country Cap:", 
                  min = -1, max = 30, value = -1)),
      mainPanel(plotOutput("Winners")))
    ),
  server = function(input,output, session) {
    observeEvent(input$region, {
      if (input$region == "Summer") {
        choices <- events %>% filter(Season == "Summer") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      } else {
        choices <- events %>% filter(Season == "Winter") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      }
    updateSelectInput(session, "nYear", choices = choices)
    })
    output$Winners = renderPlot(PropTeamWinners(input$region, input$nYear, input$sType, input$nCountry))
  } 
)
```

## The proportion of Medals Per Individual Team Members

```{r warnings = FALSE, echo = FALSE}
shinyApp(
  ui = fluidPage(
  "Choose the Olympics and Year",
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Olympics:",
                    choices = levels(SummerOrWinter), selected = "Summer"),
      selectInput("nYear", "Olympic Year:",
                  choices = c(1896), selected  = 1896),
      radioButtons("sType", "Bar Type:",
                   choices = c("Fill", "Dodge"), selected = "Fill"),
      sliderInput("nCountry", "Country Cap:", 
                  min = -1, max = 30, value = -1)),
      mainPanel(plotOutput("Winners")))
    ),
  server = function(input,output, session) {
    observeEvent(input$region, {
      if (input$region == "Summer") {
        choices <- events %>% filter(Season == "Summer") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      } else {
        choices <- events %>% filter(Season == "Winter") %>% select(Year) %>% distinct(Year) %>% arrange(Year)
      }
    updateSelectInput(session, "nYear", choices = choices)
    })
    output$Winners = renderPlot(PropTeamAthleteWinners(input$region, input$nYear, input$sType, input$nCountry))
  } 
)
```

## Sporting Trends

```{r warnings = FALSE, echo = FALSE}
shinyApp(
  ui = fluidPage(
  "Choose the Olympics and Year",
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Olympics:",
                    choices = levels(SummerOrWinter), selected = "Summer")),
      mainPanel(plotOutput("Winners")))
    ),
  server = function(input,output, session) {
    output$Winners = renderPlot(sportTotals(input$region))
  } 
)
```

# Questions

## Bibliography
